---
title: "Mining ClinicalTrials.gov with *rclinicaltrials*"
author: "by Nick Giangreco"
output: html_notebook
---

**It would be very helpful to mine current clinical trials for promising but experimental drugs. *rclinicaltrials*[https://github.com/sachsmc/rclinicaltrials] is an R package that will allow you to mine for drugs within the US clinical trials database. **

**This is an ongoing development project. Contact the author and post an issue to get it up and running!**

Setting up environment
```{r,eval=T,echo=T}
install.packages("devtools")
library(devtools)
install_github("sachsmc/rclinicaltrials")
library(rclinicaltrials)
```

I'm interested in Temozolamide-how many clinical studies are listed in the database?
```{r,eval=T,echo=T}
drug="Temozolamide"
clinicaltrials_count(query = drug)
tem<-clinicaltrials_download(query = drug)
```
This is the same number as when you enter 'Temozolamide' into the clinicaltrrials.gov search box. 


What is the status of the trials?
```{r,echo=T,eval=T}
tem<-clinicaltrials_search(query = drug)
status<-tem$status.text
status
```


What are the titles of the completed trials?
```{r,echo=T,eval=T}
tem$title[tem$status.text=="Completed"]
```