---
title: "Pubmed searching with RISMed: A Simple Example"
author: by Nick Giangreco
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Literature searches within R

**The package *RISMed* allows you to do pubmed, or any NCBI database, searches with results from R seamlessly within the R environment. This tutorial will go through a simple example of searching a term of interest in PubMed with *RISMed* functions.**

**I first set my environment and load my data,**

```{r, echo=TRUE, eval=TRUE}
#install.packages("RISmed") #if not already installed
library(RISmed)
```

**Let's say I'm interested in finding literature about Temozolamide, a common drug used to treat Glioblastoma. **

```{r,echo=TRUE,eval=TRUE}
search_query <- EUtilsSummary("Temozolamide")

print(search_query)
```
**This shows my search that will be submitted to PubMed. In this case, it looks for "Temozolamide" in all fields of PubMed submissions.**


**The summary function will yield the actual query to pubmed, based on the keyword. The EUtils wrapper function will output the query results number. Using summary(), I can see how many results I get.**
```{r,echo=TRUE,eval=TRUE}
summary(search_query)
```
**88 is exactly how many articles I would retrieve if I entered "Temozolamide[All Fields]" in the search box to PubMed.**

**Now I'd like to retrieve all PubMed articles from 2016.**

```{r,echo=TRUE,eval=TRUE}
search_query <- EUtilsSummary("Temozolamide",mindate=2016,maxdate=2016)

records<- EUtilsGet(search_query)
```

**Now I can invoke many different functions to look at these records. For example, what are the titles of the articles?**
```{r,echo=TRUE,eval=TRUE}
ArticleTitle(records)
```

**Cool, I think the first entry would be a great article to read! What's the PMID?**

```{r,echo=TRUE,eval=TRUE}
PMID(records)[1]
```

**Gee, isn't this great I can do literature searches right in my R console!? Thanks Stephanie Kovalchik of Victoria University for creating *RISMed*!**